OpReg-Boost: Learning to Accelerate Online Algorithms \\ with Operator Regression

### 2021-01-15 paper/main.tex
Third pass ++

An operator $\T: \R^n \to \R^n$ is said to be non-expansive iff $\|T \x - T \y \| \leq \|\x- \y\|$, for all $\x, \y \in \R^n$, whereas it is $\zeta$-contractive, $\zeta\in(0,1)$, iff $\|T \x - T \y \| \leq \zeta \|\x- \y\|$, for all $\x, \y \in \R^n$.

We indicate with $\prox_{\alpha g}(\cdot)$ the proximal operator of function $g$, and with $\proj_{C}(\cdot)$, the projection operator onto the set $C$.

@@

We are interested in solving the following time-varying optimization problem,
\begin{equation}\label{eq:continuous-time-problem}
	\x^*(t) \in \argmin_{\x} f(\x; t) + g(\x; t) =: F(\x; t)
\end{equation}
where $f : \R^n \times \R_+ \to \R$ is closed, proper, and $\mu$-weakly convex, whereas $g : \R^n \times \R_+ \to \R \cup \{ +\infty \}$ is closed, convex and proper function, optionally with $g \equiv 0$. Upon sampling the problem at discrete time instances $t_k$, we then obtain the sequence of time-invariant problems,
\begin{equation}\label{eq:base-problem}
	\x^*(t_k) \in \argmin_{\x} f(\x; t_k) + g(\x; t_k), \qquad k \in \N. 
\end{equation}
Our overarching goal is to set up an online algorithm of the forward-backward type,
\begin{equation}
\x_{k+1} = \prox_{\alpha g(\cdot, t_k)} (\x_k - \alpha \nabla_{\x} f(\x_k; t_k)), \quad k \in \N,
\end{equation}
that generate a sequence of approximate optimizers $\{\x_k\}_{k\in \N}$, such that the asymptotical tracking error \andrea{
\begin{equation}
\limsup_{k \to \infty} \|\x_k - \x^*(t_k)\|,
\end{equation}
is as small as possible. Our main blanket assumption is that optimizers\footnote{..} at subsequent time instances cannot be arbitrarily far apart, that is, 
\begin{equation}
\|\x^*(t_{k+1}) - \x^*(t_k)\| \leq \Delta < +\infty, \qquad k \in \N,
\end{equation}
which in turn guarantees that the path-length grows linearly in time, 
\begin{equation}
\sum_{k\in \N} \|\x^*(t_{k+1}) - \x^*(t_k)\| \leq \Delta k.
\end{equation}

Here should we formulate it in terms of functional $|f - f^*|$ instead? Let's discuss.

Somewhere we have to say that forward-backward may not converge in general, while here we ensure convergence since we are convex. This is an important point

}

\andrea{Explain a bit better what we mean by online-- I think we could be clearer?? Or maybe in the introduction.}

\andrea{Is $f$ differentiable, does it have to be?}

The first important and known result here is that if function $f$ is in $\mathcal{S}_{m,L}(\R^n)$ uniformly in $k$, then an online forward-backward algorithm can obtain linear convergence to the asymptotical error bound $\limsup_{k\to\infty}\|\x_k - \x^*(t_k)\| \leq \gamma\Delta/(1-\gamma)$, with optimal $\gamma = \frac{L-m}{L+m}$; much weaker results hold for the general convex case~\cite{Simonetto20XX}.

The second important result is that even a simple Tikhonov regularization $+ \frac{w}{2} \|x\|^2_2$, transforming the above $f \in \mathcal{S}_{m,L}(\R^n)$ to $f' \in \mathcal{S}_{m',L'}(\R^n)$ with a better $\gamma'<\gamma$, could allow for a boost in convergence and reducing the asymptotical error. In fact, if the optimizers of the true function $F = f+g$ and regularized function $F' = f'+g$ are never farther apart than $e_{\x}$, then the asymptotical error for an algorithm on $F'$ is $\gamma'\Delta'/(1-\gamma') + e_{\x}$, which can be less than the original $\gamma\Delta/(1-\gamma)$.

\andrea{say something more, change above -- these are worst case bounds..?}

@@

@@

\andrea{What happens to $\|\tv' - \tv''\|$ for $\x'$ and $\x''$ both interpolated? }

@@

\subsection{Optional auto-tuning }

\andrea{Appendix?}

@@

>>>>

An operator $\T: \R^n \to \R^n$ is said to be non-expansive iff $\|T \x - T \y \| \leq \|\x- \y\|$, for all $\x, \y \in \R^n$, whereas it is $\zeta$-contractive, $\zeta\in(0,1)$, iff $\|\T \x - \T \y \| \leq \zeta \|\x- \y\|$, for all $\x, \y \in \R^n$.

We indicate with $\prox_{\alpha g}(\cdot)$ the proximal operator of function $g$ with parameter $\alpha>0$, and with $\proj_{C}(\cdot)$, the projection operator onto the set $C$.

@@

We are interested in solving the following time-varying optimization problem,
\begin{equation}\label{eq:continuous-time-problem}
	\x^*(t) \in \argmin_{\x \in \R^n} f(\x; t) + g(\x; t) =: F(\x; t)
\end{equation}
where $f : \R^n \times \R_+ \to \R$ is closed, proper, and $\mu$-weakly convex, whereas $g : \R^n \times \R_+ \to \R \cup \{ +\infty \}$ is closed, convex and proper function, optionally with $g \equiv 0$. Upon sampling the problem at discrete time instances $t_k$, we then obtain the sequence of time-invariant problems,
\begin{equation}\label{eq:base-problem}
	\x^*(t_k) \in \argmin_{\x \in \R^n} f(\x; t_k) + g(\x; t_k), \qquad k \in \N. 
\end{equation}
The overarching goal of time-varying optimization is to set up an online algorithm $\mathcal{A}$ that generate a sequence of approximate optimizers $\{\x_k\}_{k\in \N}$, such that the asymptotical tracking error
\begin{equation}
\limsup_{k \to \infty}  \|\x_k - \x^*(t_k)\|,
\end{equation}
is as small as possible. The main blanket assumption is that there exists an optimizer trajectory $\x^*(t)$ such that,
\begin{equation}
\|\x^*(t_{k+1}) - \x^*(t_k)\| \leq \Delta < +\infty, \qquad k \in \N,
\end{equation}
which means that the optimizers can be tracked.

To fix the ideas, we focus here on online algorithms of the forward-backward type, i.e., for each $t_{k+1}$ we update the sequence $\{\x_k\}_{k \in N}$ as,
\begin{equation}\label{eq:fb}
\x_{k+1} = \prox_{\alpha g(\cdot, t_{k+1})} (\x_{k} - \alpha \nabla_{\x} f(\x_{k}; t_{k+1})), \quad k \in \N.
\end{equation}
Note that the sequence generated by~\eqref{eq:fb} could not converge in general (for a generic non-convex $f$), but it is not important for us, since we are going to regularize either the function $f$ or the algorithm $\mathcal{A}$ so to have converging sequences.

In particular, we know that for a function $f$ that is in $\mathcal{S}_{m,L}(\R^n)$ uniformly in $k$, then~\eqref{eq:fb} can obtain linear convergence to the asymptotical error bound $\limsup_{k\to\infty} \|\x^*(t_{k+1}) - \x^*(t_k)\| \leq \gamma\Delta/(1-\gamma)$, with optimal $\gamma = \frac{L-m}{L+m}$; much weaker results hold for the general convex case~\cite{Simonetto20XX}.

Another important result is that even a simple Tikhonov regularization $+ \frac{w}{2} \|x\|^2_2$, transforming the above $f \in \mathcal{S}_{m,L}(\R^n)$ to $f' \in \mathcal{S}_{m',L'}(\R^n)$ with a better $\gamma'<\gamma$, could allow for a boost in convergence and reducing the upper bound on the asymptotical error. In fact, if the optimizers of the true function $F = f+g$ and regularized function $F' = f'+g$ are never farther apart than $e_{\x}$, then the asymptotical error for an algorithm on $F'$ is $\gamma'\Delta'/(1-\gamma') + e_{\x}$, which can be less than the original $\gamma\Delta/(1-\gamma)$.

\andrea{say something more, change above ..?}

@@

\andrea{Decide $k$ or $k+1$}

@@

\andrea{What happens to $\|\tv' - \tv''\|$ for $\x'$ and $\x''$ both interpolated? Say it has to be sequential! }

@@

\begin{remark}[Important] We notice here that the convex regression and operator regression methods can be used with any starting function $f$, or operator $\mathcal{T}$. However, functions close to be in $\mathcal{S}_{m,L}(\R^n)$ (e.g., smooth $\mu$-weakly convex functions with small $\mu$) and operators close to be $\zeta$-contractive (e.g., strictly non-expansive operators) would allow for more sensible outcomes. A good point to make here is that the cost function of the regression problems~\eqref{socp} and \eqref{eq:operator-regression} measures the distance between the original function or operator to the class one is projecting them onto, thereby offering a proxy for regularization error. Moreover, one can trade-off regularization accuracy and speed tuning the free parameters $m,L$ or $\zeta$. An option for tuning $\zeta$ automatically is explored in the Appendix.
\end{remark}

@@

\subsection{Optional auto-tuning for $\zeta$ in~\eqref{eq:operator-regression}}
